---
title: "RP1"
author: "Gezhi Chen"

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.** This is a suggested formatting for your abstract. This **quarto** document is a template for your report. It includes some basic ways of working with such documents. You are invited to learn more about developing online reports with this technology. 
>
> **Keywords.** bsvars, quarto, R, Australia inflation rate

# Introduction


## The question objective, and motivation

**Objective: ** 

The purpose of this research is to explore the trend in Australia's inflation rate over the next two years.

**Question: ** 

Will Australia's inflation rate fall back to the 2-3% inflation target range in 2025, and will it reach 2.5% in 2026?

The Reserve Bank of Australia (RBA) forecasts that inflation will fall back to the target range of 2% to 3% by 2025 and reach the midpoint of this range in 2026. The purpose of the study is to forecast inflation over the next two years and assess the reasonableness of the RBA's forecast. This is because we have observed a sharp rise in inflation to 7.8% during COVID-19, and in response the RBA implemented tightening monetary policy and increased interest rates starting in late 2022. Entering mid-to-late 2023, the pace of interest rate increases has slowed down. This research will help determine whether the RBA correctly assumed that its monetary policy measures have effectively suppressed inflation and expected that inflation will fall back to the target range over the next year.

```{r Library}
#| echo: false
#| message: false
#| warning: false

library(readabs)
library(readrba)
library(xts)
library(fUnitRoots)   # ADF test - adfTest
library(tidyverse)    # for table
library(kableExtra)   # for print table
```


## Data and their properties

**Data Selection and**

Direct Inflationary Indicators:

1.$cpi_i$: Consumer Price Index (CPI): Direct measure of inflation (from ABS)

2.$gold_i$: Gold price, indicate inflationary expectations (from yahoo finance)

Economic Activity and Policy Indicators:

3.$gdp_i$: Gross Domestic Product (GDP), reflect economic growth and demand-pull inflation (from ABS)

4.$crt_i$: Cash rate target, Indicate monetary policy stance which affects inflation (from RBA)

5.$unemp_i$: Unemployment rate, indicate the level of slack in the labor market, influencing wage-push inflation (from ABS)

Market and External Trade Indicators:

6.$nloan_i$: New loan commitments, reflect credit growth which can fuel economic activity and inflation (from ABS)

7.$export_i$: International exports, affect trade balance and currency strength, influencing imported inflation (from ABS)

8.$import_i$: International imports: Directly affect inflation through the cost of imported goods (from ABS)

9.$aord_i$: All Ordinaries Index (AORD), reflects investor confidence and economic activity which can be pre-emptive indicators of inflation (from yahoo finance)

10.$exr_i$: AUD/USD exchange rate, affects the price of imports and exports, contributing to inflation (from yahoo finance)


```{r Data downloading, fig.pos="H"}
#| echo: false
#| message: false
#| warning: false

### Data downloading

# 1.Inflation / CPI 
# 6401.0 Consumer Price Index, Australia
# series_id = "A2325846C": Index Numbers ;  All groups CPI ;  Australia ;
cpi_download  = read_abs(series_id = "A2325846C")     
cpi_data      = xts(cpi_download$value, cpi_download$date)

# 2.Gold price
gold_link       = "https://query1.finance.yahoo.com/v7/finance/download/GC%3DF?period1=1262304000&period2=1703980800&interval=1mo&filter=history&frequency=1mo&includeAdjustedClose=true"
gold_download   = read.csv(gold_link)
gold_data       = gold_download[,6]
gold_data       = data.frame(gold_download[,1], gold_data)
colnames(gold_data) = c('date', 'gol')
gold_data$date  = as.Date(as.character(gold_data$date),format="%Y-%m-%d") 
gold_data       = xts(gold_data$gol, gold_data$date)
gold_data       = apply.quarterly(gold_data, mean)

# 3.GDP
# 5206.0 Australian National Accounts: National Income, Expenditure and Product
# series_id = "A2304370T": Gross domestic product: Chain volume measures - Percentage changes
gdp_download  = read_abs(series_id = "A2304370T")     
gdp_data      = xts(cpi_download$value, cpi_download$date)

# 4.Cash rate target

crt_download   = read_rba(series_id = "FIRMMCRTD")   
crt_data       = xts(crt_download$value, crt_download$date)
crt_data       = apply.quarterly(crt_data, mean)     # change daily to quarterly
crt_data       = xts(crt_data, seq(as.Date("1990-03-01"), by = "quarter", length.out = length(crt_data)))

# 5.Unemployment rate
# 6202.0 Labour Force, Australia
# series_id = "A84423050A": Unemployment rate ;  Persons ; seasonal adjust
unemp_download = read_abs(series_id = "A84423050A")     
unemp_data     = xts(unemp_download$value, unemp_download$date)
unemp_data     = apply.quarterly(unemp_data, mean) # change daily to quarterly
unemp_data     = xts(unemp_data, seq(as.Date("1978-03-01"), by = "quarter", length.out = length(unemp_data)))

# 6.New loan
# 5601.0 Lending Indicators
# series_id = "A108296973X"： Households ;  Housing Finance ;  Total housing excluding refinancing ;  New loan commitments ;  Value ; seasonal adjust
nloan_download   = read_abs(series_id = "A108296973X")     
nloan_data       = xts(nloan_download$value, nloan_download$date)
nloan_data       = apply.quarterly(nloan_data, mean)

# 7.Export
# 5368.0 International Trade in Goods
# series_id = "A2718603V": Debits, Total goods ;
export_download = read_abs(series_id = "A2718603V")     
export_data     = xts(export_download$value, export_download$date)
export_data     = abs(export_data)
export_data     = apply.quarterly(export_data, mean)
 
# 8.Import
# 5368.0 International Trade in Goods
# series_id = "A2718577A": Credits, Total goods ;
import_download  = read_abs(series_id = "A2718577A")     
import_data      = xts(import_download$value, import_download$date)
import_data      = apply.quarterly(import_data, mean)

# 9.AORD
aord_link       = "https://query1.finance.yahoo.com/v7/finance/download/%5EAORD?period1=1262304000&period2=1703980800&interval=1mo&filter=history&frequency=1mo&includeAdjustedClose=true"
aord_download   = read.csv(aord_link)
aord_data       = aord_download[,6]
aord_data       = data.frame(aord_download[,1], aord_data)
colnames(aord_data) = c('date', 'aord')
aord_data$date  = as.Date(as.character(aord_data$date),format="%Y-%m-%d") 
aord_data       = xts(aord_data$aord, aord_data$date)
aord_data       = apply.quarterly(aord_data, mean)

# 10. AUD/USD
exr_link        = "https://query1.finance.yahoo.com/v7/finance/download/AUDUSD%3DX?period1=1262304000&period2=1703980800&interval=1mo&filter=history&frequency=1mo&includeAdjustedClose=true"
exr_download    = read.csv(exr_link)
exr_data        = exr_download[,6]
exr_data        = data.frame(exr_download[,1], exr_data)
colnames(exr_data) = c('date', 'exr')
exr_data$date   = as.Date(as.character(exr_data$date),format="%Y-%m-%d") 
exr_data        = xts(exr_data$exr, exr_data$date)
exr_data        = apply.quarterly(exr_data, mean)


```

```{r Data combine, fig.pos="H"}
#| echo: false
#| message: false
#| warning: false

### Data plot
# All Variables
all_data             = na.omit(merge(cpi_data,    gold_data, 
                                     gdp_data,    crt_data, 
                                     unemp_data,  nloan_data,  
                                     export_data, import_data, 
                                     aord_data,   exr_data ))
colnames(all_data)   = c("cpi_data",    "gold_data", 
                         "gdp_data",    "crt_data", 
                         "unemp_data",  "nloan_data", 
                         "export_data", "import_data",
                         "aord_data",   "exr_data")
```


```{r Data plot, fig.align='center',fig.pos='H'}
#| echo: false
#| message: false
#| warning: false

## Line plot
par(mfcol = c(5, 2), mar = c(2, 2, 2, 2))

for (i in 1:10) {
  ts.plot(all_data[, i], main = colnames(all_data)[i], 
          ylab = "", xlab = "", col = "darkblue")
}
```


```{r ACF plot, fig.align='center',fig.pos='H'}
#| echo: false
#| message: false
#| warning: false

## ACF plot
par(mfcol = c(5, 2), mar=c(2,2,2,2))
for (i in 1:10){
  acf = acf(all_data[,i], plot = FALSE)[1:20]
  plot(acf, main = "")
  title(main = paste(colnames(all_data)[i]), line = 0.5)
}
```

```{r ACF test, fig.align='center',fig.pos='H'}
#| echo: false
#| message: false
#| warning: false

## AR

# find the optimal lag 
ol.cpi.aic.ar    = (ar(cpi_data, order.max=20, aic=TRUE, method="ols"))
ol.gold.aic.ar   = (ar(gold_data, order.max=20, aic=TRUE, method="ols"))
ol.gdp.aic.ar    = (ar(gdp_data, order.max=20, aic=TRUE, method="ols"))
ol.crt.aic.ar    = (ar(crt_data, order.max=20, aic=TRUE, method="ols"))
ol.unemp.aic.ar  = (ar(unemp_data, order.max=20, aic=TRUE, method="ols"))
ol.nloan.aic.ar  = (ar(nloan_data, order.max=20, aic=TRUE, method="ols"))
ol.export.aic.ar = (ar(export_data, order.max=20, aic=TRUE, method="ols"))
ol.import.aic.ar = (ar(import_data, order.max=20, aic=TRUE, method="ols"))
ol.aord.aic.ar   = (ar(aord_data, order.max=20, aic=TRUE, method="ols"))
ol.exr.aic.ar    = (ar(exr_data, order.max=20, aic=TRUE, method="ols"))


## ADF test

# ol.cpi.aic.ar$order
adf.cpi   = adfTest(cpi_data, lags=5, type="c")              # don't reject -> non-stationary
dadf.cpi  = adfTest(diff(cpi_data), lags=4, type="nc")        # don't reject -> non-stationary
d2adf.cpi = adfTest(diff(diff(cpi_data)), lags=3, type="nc")  # reject -> (I2 is stationary)
# adf.cpi@test$p.value
#-> integration order = 2

# ol.gold.aic.ar$order
adf.gold   = adfTest(gold_data, lags=16, type="c")               # don't reject -> non-stationary
dadf.gold  = adfTest(diff(gold_data), lags=15, type="nc")        # don't reject -> non-stationary
d2adf.gold = adfTest(diff(diff(gold_data)), lags=14, type="nc")  # reject -> (I2 is stationary)
# adf.gold@test$p.value
#-> integration order = 2

# ol.gdp.aic.ar$order
adf.gdp   = adfTest(gdp_data, lags=5, type="c")              # don't reject -> non-stationary
dadf.gdp  = adfTest(diff(gdp_data), lags=4, type="nc")       # don't reject -> non-stationary
d2adf.gdp = adfTest(diff(diff(gdp_data)), lags=3, type="nc")  # reject -> (I2 is stationary)
# adf.gdp@test$p.value
#-> integration order = 2

# ol.crt.aic.ar$order
adf.crt  = adfTest(crt_data, lags=20, type="c")               # don't reject -> non-stationary
dadf.crt = adfTest(diff(crt_data), lags=19, type="nc")        # reject -> (I1 is stationary)
# adf.crt@test$p.value
#-> integration order = 1

# ol.unemp.aic.ar$order
adf.unemp  = adfTest(unemp_data, lags=3, type="c")           # don't reject -> non-stationary
dadf.unemp = adfTest(diff(unemp_data), lags=2, type="nc")    # reject -> (I1 is stationary)
# adf.unemp@test$p.value
#-> integration order = 1

# ol.nloan.aic.ar$order
adf.nloan  = adfTest(nloan_data, lags=2, type="c")           # don't reject -> non-stationary
dadf.nloan = adfTest(diff(nloan_data), lags=1, type="nc")    # reject -> (I1 is stationary)
# adf.nloan@test$p.value
#-> integration order = 1

# ol.export.aic.ar$order
adf.export  = adfTest(export_data, lags=13, type="c")           # don't reject -> non-stationary
dadf.export = adfTest(diff(export_data), lags=12, type="nc")    # reject -> (I1 is stationary)
# adf.export@test$p.value
#-> integration order = 1

# ol.import.aic.ar$order
adf.import  = adfTest(import_data, lags=12, type="c")           # don't reject -> non-stationary
dadf.import = adfTest(diff(import_data), lags=11, type="nc")    # reject -> (I1 is stationary)
# adf.import@test$p.value
#-> integration order = 1

# ol.aord.aic.ar$order
adf.aord   = adfTest(aord_data, lags=15, type="c")               # don't reject -> non-stationary
dadf.aord  = adfTest(diff(aord_data), lags=14, type="nc")        # don't reject -> non-stationary
d2adf.aord = adfTest(diff(diff(aord_data)), lags=13, type="nc")  # reject -> (I2 is stationary)
# adf.aord@test$p.value
#-> integration order = 2

# ol.exr.aic.ar$order
adf.exr   = adfTest(exr_data, lags=20, type="c")               # don't reject -> non-stationary
dadf.exr  = adfTest(diff(exr_data), lags=19, type="nc")        # don't reject -> non-stationary
d2adf.exr = adfTest(diff(diff(exr_data)), lags=18, type="nc")  # reject -> (I2 is stationary)
# adf.exr@test$p.value
# #-> integration order = 2

Unit_Root_Test_table <- 
  tibble( " " = c("cpi", "gold", "gdp", "crt", "unemp", "nloan", "export","import", "aord", "exr"),
          "p value of ADF test of AR" 
          = round(c(adf.cpi@test$p.value,    adf.gold@test$p.value,
                    adf.gdp@test$p.value,    adf.crt@test$p.value,    
                    adf.unemp@test$p.value,  adf.nloan@test$p.value,
                    adf.export@test$p.value, adf.import@test$p.value,
                    adf.aord@test$p.value,   adf.exr@test$p.value),4),
          "p value of ADF test of diff-AR" 
          = round(c(dadf.cpi@test$p.value,    dadf.gold@test$p.value,
                    dadf.gdp@test$p.value,    dadf.crt@test$p.value,    
                    dadf.unemp@test$p.value,  dadf.nloan@test$p.value,
                    dadf.export@test$p.value, dadf.import@test$p.value,   
                    dadf.aord@test$p.value,   dadf.exr@test$p.value),4),
          "p value of ADF test of diff-diff-AR" 
          = round(c(d2adf.cpi@test$p.value,    d2adf.gold@test$p.value,
                    d2adf.gdp@test$p.value,
                    NA, NA, NA, NA, NA,
                    d2adf.aord@test$p.value,   d2adf.exr@test$p.value),4),
          "conclusion" 
          = c("cpi~I(2)",    "gold~I(2)", 
              "gdp~I(2)",    "crt~I(1)",    
              "unemp~I(1)",  "nloan~I(1)", 
              "export~I(1)", "import~I(1)", 
              "aord~I(2)",   "exr~I(2)"
          ),
  )


kable(Unit_Root_Test_table, align = "c") %>% 
  kable_styling(font_size = 8, 
                fixed_thead = TRUE, 
                full_width = FALSE, 
                position = "center",
                latex_options = c("HOLD_position"),
                bootstrap_options = c("striped", "hover", "bordered", "responsive", "dark"))

```

## The model and hypothesis







## References {.unnumbered}